---
title: "Silvestrum Assignment"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
---

#Load the libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(pwr)
```


Set the File path for Data extraction
```{r}
#Setting my filepaths
rootdir <- ("/Users/javipatron/Documents/MEDS/Courses/eds222/homework/eds222-finalproject")
data <- file.path(rootdir,"data")

```

Reading the .csv files
```{r}
#Load the shared Mangrove Data from Silvestrum
mangrove_df_a <- read_csv(file.path(data,"clean_monitoring_data_javier.csv")) |> clean_names() #Clean names will make the column names as lower case

mangrove_df_b <- read_csv(file.path(data, "monitoring_siteB_data_javier.csv")) |> clean_names() |> na.omit() #homogenizing the column names and adding NA to values

```


# Site A

Understand the data with R functions
```{r}
#Wranggle the data for table A
dim(mangrove_df_a) #Dimentions
names(mangrove_df_a) #Column Names
skimr::skim(mangrove_df_a) #General Skim through the data set. Column per column results
summary(mangrove_df_a) #Stats analysis per column

unique(mangrove_df_a$plantation_year) #Get to know the different years
unique(mangrove_df_a$plot_size_m2) # Get to know the unique plot sizes


```


#Cleaning and adding some changes to the data
```{r}
#Changing some data
mangrove_df_a$plantation_year <- as.character(mangrove_df_a$plantation_year) # Making the years as character instead of numbers
mangrove_df_a$plot <- as.character(mangrove_df_a$plot) # Making the plot names as character
mangrove_df_a$plot_size_m2 <- as.factor(mangrove_df_a$plot_size_m2) # Making the plot sizes as factor instead of numeric

```

Create a table that is grouped by the each plantation year containing the total number of sample plots and the total number of samples trees per plantation.

```{r}

mangrove_summary <- mangrove_df_a |> 
  group_by(plantation_year) |> 
  summarise(plot_count = length(unique(plot)),
          tot_samples = n(),
          "mean_carbon (Mg/ha)" = round(mean(total_tree_mg_c_ha),3))
          
knitr::kable(mangrove_summary)

```

Create a table that shows all the samples per each plot
```{r}

plots_a <- mangrove_df_a |> 
  group_by(plantation_year,plot,plot_size_m2) |> 
  summarise(plot = unique(plot),
            sample_count = n(),
          "mean_carbon (Mg/ha)" = round(mean(total_tree_mg_c_ha),3))

```

Create a table of samples per plot but grouping by sample size
```{r}
mangrove_plots_summary <- mangrove_df_a |>
  filter(plot_size_m2 == 153.938 ) |> 
  group_by(plantation_year) |> 
  summarise(plot_size_m2 = unique(plot_size_m2),
            plot_count = length(unique(plot)),
            tot_samples = n(),
            mean_plant_size_cm = (mean(cd_chatting_m)),
            mean_tot_carbon_g_c = (mean(total_tree_kg_c)),
            sd_plant_size = sd(cd_chatting_m),
            sd_tot_carbon_kg_c = sd(total_tree_kg_c))

kable(mangrove_plots_summary)
```


# Site B

Create a table that summarises the site B by year
```{r}
mangrove_plots_b_summary <- mangrove_df_b |> 
  group_by(year) |> 
  summarise(plot_count = length(unique(name)),
            tot_samples = n(),
            mean_total_c_t_ha = (mean(total_c_t_ha)),
            sd_total_c_t_ha = sd(total_c_t_ha))
```

Create a table that shows all the plot samples 
```{r}
plots_b <- mangrove_df_b |> 
  group_by(year, name) |> 
  summarise(sample_count = n(),
          "mean_carbon (Mg/ha)" = round(mean(total_c_t_ha),6))
```

# Joining both dataframes with plot sizes of 154m2 only
```{r}
#Cleaning and getting the site A ready for the stacking
tidy_a <- mangrove_df_a |> 
  filter(plot_size_m2 == 153.938) |> 
  select(plantation_year, plot, height_cm,total_tree_kg_c, total_tree_mg_c_ha) |> 
  mutate(site = "A") |> 
  relocate(site, .before = plantation_year) |> 
  rename(year = plantation_year,
         plot_name = plot,
         total_tree_c_kg = total_tree_kg_c,
         "total_tree_c_Mg_ha" = total_tree_mg_c_ha)

```

```{r}
# Cleaning and getting the site B for the stacking with site A
tidy_b <- mangrove_df_b |> 
  select(year, plot, height_cm, total_c_kg, total_c_t_ha) |> 
  rename(plot_name = plot,
         total_tree_c_kg = total_c_kg,
         "total_tree_c_Mg_ha" = total_c_t_ha) |> 
    mutate(site = "B") |> 
    relocate(site, .before = year)

```


```{r}
# Creating a row binding to have the same columns but all the information
full_df <- rbind(tidy_a, tidy_b)

summary_full_df <- full_df |> 
  group_by(site, year) |> 
  summarise(plot_count = length(unique(plot_name)),
            tot_samples = n(),
            mean_tree_height_cm = round(mean(height_cm),2),
            mean_tot_carbon_Mg_ha = round(mean(total_tree_c_Mg_ha),2),
            sd_plant_size = round(sd(height_cm),2),
            sd_tot_carbon_Mg_ha = round(sd(total_tree_c_Mg_ha),2))
```

Adding the A/R Methodological Tool

$ n = \frac{N  \text{ }*  \text{ }tvalue^2  \text{ }* \text{ }(\epsilon\ w * s)^2)}{N  \text{ }*\text{ } E^2 + \text{ }tvalue^2  \text{ }* \text{ }\epsilon\ w * s^2} $


where:
*n* = Number of sample plots required for estimation of biomass stocks within the project
boundary; dimensionless.
*N* = Total number of possible sample plots within the project boundary (i.e. the sampling
space or the population); dimensionless.
*tvalue* = Two-sided Student´s t-value, at infinite degrees of freedom, for the required confidence level; dimensionless. (Table -> 90% = 1.645)
*w* = Relative weight of the area of stratum i (i.e. the area of the stratum i divided by the project.
*E* = Acceptable margin of error (i.e. one-half the confidence interval) in estimation of biomass stock within the project boundary; t d.m. (or t d.m. ha-1).


```{r}

sample_plots <- function(N, tvalue, w, s, E){
  
  epsilon = for (i in range(table)){
    w * s
  }
  n <- (N * tvalue^2 * sum(epsilon)^2) / (N * E^2 + tvalue ^2 * sum(epsilon)^2)
  
  return(n)
}

```




